---
title: "C9W3 Assignment"
author: "Nicholas Neo"
date: "13 July 2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Chicken Rice in Singapore

An R markdown presentation that shows the top 8 famous Chicken Rice stalls in Singapore for the tourists.  
[Source](https://www.womensweekly.com.sg/gallery/family/best-chicken-rice-stalls-we-love-in-singapore/)


## Best 8 Chicken Rice stalls in Singapore

- Jew Kit Hainanese Chicken Rice
- Boneless Chicken Rice Katong Shopping Centre
- Boon Tong Kee Chicken Rice
- Tong Fong Fatt Hainanese Boneless Chicken Rice
- Poh Poh Roasted Chicken Rice
- Hainanese Delicacy
- Five Star Hainanese Chicken Rice
- Ming Kee Chicken Rice

## Venue on the map

```{r plot_venue, echo=FALSE, warning=FALSE}
library(leaflet)
coords <- data.frame(lat = c(1.299950, 1.303940, 1.343340, 1.313170, 1.320520, 1.307180, 1.307200, 1.348290),
           lng = c(103.840200, 103.901350, 103.775140, 103.785640, 103.742630, 103.833810, 103.906420, 103.853170))
popup <- c("Jew Kit Hainanese Chicken Rice", "Boneless Chicken Rice Katong Shopping Centre", "Boon Tong Kee Chicken Rice", "Tong Fong Fatt Hainanese Boneless Chicken Rice", "Poh Poh Roasted Chicken Rice", "Hainanese Delicacy", "Five Star Hainanese Chicken Rice", "Ming Kee Chicken Rice")
coords %>% 
    leaflet() %>%
    addTiles() %>%
    addMarkers(popup = popup)
```

