---
title: "C9W4 Assignment (Reproducible Pitch)"
author: "Nicholas Neo"
date: "13 July 2020"
output: ioslides_presentation
---

## About the Dataset
- Diamonds dataset from the ggplot2 package.
- Consists of 53940 diamonds with 10 characteristics
**Loading the diamonds dataset**
```{r, warning = FALSE}
library(ggplot2)
head(diamonds)
```

## Problem Statement
- Using the following characteristics (cut, color, clarity and carat) of the diamond to predict the price of the diamond.
- General linear model is used to fit the model and use it for prediction.
- Plot is made with the Price of the diamond (y-axis) against the carat of the diamond (x-axis) using **ggplot2**.
- Deployed using shiny.io web app.

## How to use the app?
- From the sidebar, choose your diamond by selecting the cut, color, clarity, and carat. 
- The general linear model will predict the price of the diamond based on your inputs, and displays it at the bottom of the main panel. 
```{r, eval=FALSE, warning = FALSE}
pred <- predict(fit, newdata = data.frame(carat = input$car,
                                              cut = input$cut,
                                              color = input$col,
                                              clarity = input$clar))
```
- The app will then plot the regression line (in blue).
```{r, eval=FALSE}
plot <- ggplot(data=diam, aes(x=carat, y = price))+
      ...  # remaining code to get the plot
    plot
```

## Your Turn!
- Try out the [Diamond Predictor](https://thatnicneo.shinyapps.io/c9w4assignment/) now!
- Source Code:   
[ui](https://github.com/nicholasneo78/JHU-DataScience-Assignment/blob/master/Course%209.4/C9W4Assignment/ui.R)  
[server](https://github.com/nicholasneo78/JHU-DataScience-Assignment/blob/master/Course%209.4/C9W4Assignment/server.R)